The evolution of gene expression in the term placenta of viviparous mammals

### 2015-12-20 manuscript/manuscript.Rnw
fix up bonobo discussion

The top 50 expressed genes as measured by \ac{FPKM} in the term
placenta of the bonobo are shown in \cref{tab:top50_bonobo_placenta}
with a list of all genes in Supplemental
\cref{tab:all_pan_paniscus_placenta}. Expression in the bonobo
placenta differs somewhat from that of the human, as evidenced by the
makeup of the top 50 genes (\cref{tab:top50_bonobo_placenta}). The highest expression
(\ac{FPKM}=%
\Sexpr{gene.expression("pan paniscus","PRG2")}%
%
) in bonobo is \textit{PRG2}, which is involved in insulin
growth factor signaling and angiogenesis during pregnancy \cite{Weyer.ea2011:Placentalregulationofpeptide}. \textit{PRG2} is also expressed at relatively high levels in the
human placenta (\ac{FPKM}=\Sexpr{gene.expression("homo sapiens","PRG2")}; Supplemental \cref{tab:all_homo_sapiens_placenta}), although at
comparatively much lower levels than the bonobo. A bonobo gene closely
related to the chorionic somatomammotropins (Q95MK7\_PANTR) is also
expressed at high levels (\ac{FPKM}=1232.56; \cref{tab:top50_bonobo_placenta}). Two other Pan genes
included within this family (ENSPTRG00000030805 which shows similarity
to GH2 and ENSPTRG00000034560) show expression levels of \ac{FPKM}>89. Six
pregnancy-specific beta-1-glycoprotein genes were expressed in the
bonobo placenta with \ac{FPKM} ranging from 20.885-167.961 (Supplemental
Table X). Placental expression of the genes involved in estrogen and
progesterone production (CYP19A1, HSD3B1 respectively) show lower
expression in the bonobo as compared to humans (CYP19A1 \ac{FPKM}= 48.6877;
HSD3B1 \ac{FPKM}= 54.5301; Table 8). In addition, CGA is expressed in the
bonobo placenta (\ac{FPKM}=202.629), but at much lower levels than humans
as well (Table 8). Two Ensembl-annotated beta subunits of the
chorionic gonadotropin genes are expressed in bonobo, one related to
CGB and another related to CGB7; both are not highly expressed (Table
8). Among the other top 50 expressed genes, there is high expression
of ribosomal proteins (e.g., RPL18, RPL23), keratins (e.g., KRT7,
KRT19), heat-shock proteins (e.g., HSPA5, HSPA1A), serpins (SERPINE1,
SERPINE2), insulin growth factor 2 (IGF2) and the hormone
adrenomedullin (ADM).

>>>>

The top 50 expressed genes as measured by \ac{FPKM} in the term
placenta of the \bonobo are shown in \cref{tab:top50_bonobo_placenta}
with a list of all genes in Supplemental
\cref{tab:all_pan_paniscus_placenta}. Expression in the \bonobo
placenta differs somewhat from that of the \human, as evidenced by the
makeup of the top 50 genes (\cref{tab:top50_bonobo_placenta}). The highest expression
 in \bonobo is %
\Sexpr{gene.expression("pan paniscus","PRG2",complete=TRUE)}, %
%
 which is involved in insulin
growth factor signaling and angiogenesis during pregnancy \cite{Weyer.ea2011:Placentalregulationofpeptide}. \textit{PRG2} is also expressed at relatively high levels in the
human placenta (\ac{FPKM}=\Sexpr{gene.expression("homo sapiens","PRG2")}; Supplemental \cref{tab:all_homo_sapiens_placenta}), although at
comparatively much lower levels than the \bonobo. A \bonobo gene closely
related to the chorionic somatomammotropins (Q95MK7\_PANTR) is also
expressed at high levels (\ac{FPKM}=1232.56; \cref{tab:top50_bonobo_placenta}). Two other Pan genes
included within this family (ENSPTRG00000030805 which shows similarity
to GH2 and ENSPTRG00000034560) show expression levels of \ac{FPKM}>89. Six
pregnancy-specific beta-1-glycoprotein genes were expressed in the
\bonobo placenta with \ac{FPKM} ranging from 20.885-167.961 (Supplemental
\cref{tab:all_pan_paniscus_placenta}). Placental expression of the genes involved in estrogen and
progesterone production (CYP19A1, HSD3B1 respectively) show lower
expression in the bonobo as compared to humans (CYP19A1 \ac{FPKM}= 48.6877;
HSD3B1 \ac{FPKM}= 54.5301; Table 8). In addition, CGA is expressed in the
bonobo placenta (\ac{FPKM}=202.629), but at much lower levels than humans
as well (Table 8). Two Ensembl-annotated beta subunits of the
chorionic gonadotropin genes are expressed in bonobo, one related to
CGB and another related to CGB7; both are not highly expressed (Table
8). Among the other top 50 expressed genes, there is high expression
of ribosomal proteins (e.g., RPL18, RPL23), keratins (e.g., KRT7,
KRT19), heat-shock proteins (e.g., HSPA5, HSPA1A), serpins (SERPINE1,
SERPINE2), insulin growth factor 2 (IGF2) and the hormone
adrenomedullin (ADM).
